---
title: "Tidyverse Solo Exercise"
output: html_notebook
---
```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
```

Read in districts.csv into a tibble named districts. How many rows and columns does it contain?
```{r}
districts <- read.csv(file = "data/districts.csv")
x <- dim(districts)[1]
y <- dim(districts)[2]
sprintf("The district tibble has %d rows and %d columns.", x, y)
```
Notice that the first row corresponds to the whole State of Tennessee. Remove this row and save the result back to districts.
```{r}
districts <- districts[-1, ]
```


How many districts have a proficiency rate of at least 80% for both alg_1 and eng_1?
```{r}
dis_80 <- length(which(districts$alg_1 >= 80 & districts$eng_1 >= 80))
sprintf("%d districts have a proficiency rate of at least 80%% for both Algebra 1 and English 1", dis_80)
```
How many districts have a proficiency rate less than 50% for either alg_1 or eng_1?
```{r}
dis_50 <- length(which(districts$alg_1 < 50 | districts$eng_1 < 50))
sprintf("%d districts have a proficiency rate of less than 50%% for either Algebra 1 or English 1", dis_50)
```
Which district has the lowest graduation rate?

```{r}
districts[which.min(districts$grad), c("system_name", "grad")]
```
Within the Mid Cumberland region, which district has the highest ACT composite?
```{r}
districts_mid_cumberland <- districts %>%
                              filter(region == "Mid Cumberland")
districts_mid_cumberland[
                        which.max(districts_mid_cumberland$act_composite),
                        c("system_name", "act_composite")
                        ]
```
Create a histogram showing the distribution of graduation rates. What can you say about this distribution?

As you can see the distribution is skewed to the left with a few outliers in the data. The median value for graduation rates across districts is 91.5%


```{r}
districts_grad <- districts %>%
                    drop_na(grad)#Drop all NA rows for graduation rates
```

```{r}
ggplot(districts_grad, aes(x = grad))+
  geom_histogram(color = "black",
                 fill = "#00fd72",
                 bins = 50) +
  geom_vline(aes(xintercept = median(grad)),
             color = "red",
             linetype = "dashed",
             linewidth = 1) +
  annotate("text",
           x = 85,
           y = 29,
           label = paste("Median\n= ", median(districts_grad$grad), "%"),
           color = "red") +
  labs(title = "Distribution of Graduation Rates in Tennessee School Districts",
       x = "Graduation Rate %",
       y = "Count")
```
Create a scatter plot to compare alg_1 proficiency rates to alg_2 rates. What do you notice? Facet this plot by region. Does anything stand out when you facet the plots?

Although stronger in some regions than in others, there seems to be a constant, positive correlation between  the Algebra 1 and Algebra 2 proficiency rates.
```{r}
districts_alg_prof <- districts %>%
                        drop_na(alg_1, alg_2)
                        #Drop NA rows only in alg_1 and alg_2 columns
```


```{r}                      
ggplot(districts_alg_prof, aes(x = alg_1, y = alg_2))+
  geom_point()+
  facet_wrap(~ region)+
  labs(title = "Algebra 2 vs. Algebra 1 Proficiency Rates for each District in TN by Region",
       x = "Algebra 1 Prof. Rates",
       y = "Algebra 2 Prof. Rates")
```

Create a bar chart showing the total enrollment by region. Which region has the highest total enrollment? Which has the smallest?

From the graph we can see that the Mid Cumberland region has the most number of students with 275916. The region with the least amount of students is Northwest with 38682. 

```{r}
district_enrollment_by_region <- districts %>%
                                    group_by(region) %>%
                                    drop_na(enrollment) %>% 
                                    summarize(n = sum(enrollment))
```

```{r}
ggplot(district_enrollment_by_region, aes(x = region, y = n, fill = region))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = n), vjust = 1.6, color = "white", size = 3.5)+
  labs(title = "Total Enrollment in TN Scools by Region", y = "Enrollment #")+
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```
When creating this bar chart you may have noticed that some districts have missing enrollment values. For how many districts is this the case?
```{r}
missing_enr <- sum(is.na(districts$enrollment))
sprintf("There are %d districts that have missing data for enrollment.",missing_enr)
```


What is the mean graduation rate across all districts? What might be wrong with using just the regular mean to assess average graduation rates?

```{r}
avg_grad_rate <- mean(districts$grad, na.rm = TRUE)
sprintf("The average graduation rate across all districts is %.1f%%", avg_grad_rate)
```
The problem with using regular mean function is that each of the districts has a different number of enrolled students. We can account for this using a weighted average. 

Redo the previous question but use a weighted average (weighted.mean) graduation across all districts, weighing by enrollment. How much does this change your answer? Can you explain using the data the reason for the big change from using the mean?

```{r}
dis_weight <- districts %>% 
                select(system_name, enrollment, grad) %>% 
                drop_na(enrollment, grad) %>% 
                mutate(enr_weight_factor = enrollment/sum(enrollment))#Weighting Factor will be the proportion of enrollment to the total number enrolled in TN

avg_grad_rate_weighted <- weighted.mean(dis_weight$grad, dis_weight$enr_weight_factor, na.rm = TRUE)
sprintf("The weighted average graduation rate by enrollment across all districts is %.1f%%", avg_grad_rate_weighted)
```
```{r}
dis_weight
```
As you can see in the table above, each district has a different number of enrolled students. To account for this when calculating the average graduation rate in TN, we take a proportion of enrollment to the total number of enrolled students in TN for each district. This then gives a weight factor that can be used to properly weight each of the graduation rates in each district. After being multiplied by the weight factor, we can add them up to get a value of 87.3% weighted average graduation rate.  



Create a boxplot showing graduation rates per region. Does anything stand out?

Most of the school districts have meadian graduation rates greater than 80%. There is one extreme outlier in the Mid cumberland region at about 10%. 
```{r}
districts %>% 
  drop_na(grad) %>% 
  ggplot(aes(x = region, y = grad, color = region))+
    geom_boxplot(outlier.colour = "red", outlier.shape = 8)+
    labs(title = "Graduation Rate TN Scools by Region", y = "Garadaution Rate %")+
    theme(axis.text.x = element_text(angle = 45, hjust=1))
  
```

Find the weighted average of graduation rates by region using enrollment as weights. Compare the results you get for the weighted average to what you see from the boxplots.

```{r}
districts %>% 
  select(grad, enrollment, region) %>% 
  drop_na() %>% 
  mutate(enr_weight = enrollment/sum(enrollment))%>%
  group_by(region) %>% 
  summarize(weighted_mean_grad = weighted.mean(grad, enr_weight), med_grad = median(grad))
```

For many districts, values for alg_2 are lower than for alg_1. Create a histogram showing the distribution of differences (alg_1 - alg_2).

```{r}
diff_alg <- districts %>% 
              select(system_name, alg_1, alg_2) %>% 
              mutate(diff = alg_1 - alg_2)

ggplot(diff_alg, aes(x = diff))+
  geom_histogram(na.rm = TRUE)+
  labs(title = "Distribution of Algebra 1 to Algebra 2 Proficiency Rate Differences across Districts")+
  xlab("Difference Value (%)")
```

Which school had the largest drop from alg_1 to alg_2?

```{r}
diff_alg[which.max(diff_alg$diff), ]
```



 For what percentage of schools is it true that alg_2 is larger than alg_1? Is there a similar drop off for eng_2 and eng_3?

```{r}
greater_alg_2 <- diff_alg %>% 
                    count(diff>0) %>% 
                    drop_na()
                    

18/99
```
```{r}
districts %>% 
  mutate(diff_eng = eng_2 - eng_3) %>% 
  select(system_name, diff_eng)
```


You may have noticed that a lot of rows are missing values. Which district has the largest number of missing values? What do you notice about schools that have a lot of missing values?
```{r}
districts %>% 
  mutate(nancount = rowSums(is.na(.))) %>% 
  filter(nancount >= 5) %>% 
  arrange(desc(nancount))

  

```
Find the correlation between graduation rate and all other variables. Create a horizontal bar chart showing these correlations. Make sure that your plot is ordered by correlation values. What do you notice from these correlations?
```{r}


```














